---
title: "A general framework to run analyses in Dias et al.xxxxx"
author:
- Gabriel Santos
- Guilherme Sanches
date: "last time knit: `r format(Sys.time(),  '%Y-%m-%d (%H:%M:%S)')`"
output:
  html_document:
    theme: paper
    code_folding: hide
    toc: true
    toc_float:
      collapsed: true
    cache: false
---

```{r, echo=FALSE, cache=FALSE}
knitr::opts_chunk$set(out.width=14,digits = 3)
options(scipen = 1, digits = 3) 
## numbers >= 10^5 will be denoted in scientific notation,
## and rounded to 2 digits

```

# Readme

# Setting & SessionInfo

Define packages used in this paper

```{r message=FALSE, warning=FALSE, result='hide'}
library(openxlsx)
library(car)
library(performance)
library(tidyverse)
library(MuMIn)
library(performance)
library(lme4)
library(see)
library(DHARMa)
library(sjPlot)
library(ggplot2)
library(ggeffects)

options(na.action=na.fail)	#Define na action to run dredge function from MuMIn
```

Show full information about R, packages and its dependences

```{r}
sessionInfo()
```

# Load data {.tabset}

<!-- 
#setwd(choose.dir())    #Usar esse Guilherme!
#setwd("C:/Artigos e resumos publicados submetidos ideias/Notes/1 - Projetos/01 - Colaborações em projetos e artigos em desenvolvimento/Projetos Jeronimo - Fiocruz")
#dir("./Dados e scripts")
#dados<-read.xlsx("./Dados e scripts/Book1.xlsx",sheet="Goias_CLEAN_2")
#dados_eclosion<-read.xlsx("./Dados e scripts/Experimento 1 (Temperatura).xlsx",sheet="Ovitrampa Goiás ")   #Tem um "espaço a mais na aba ovitrampa Goiais, cuidado!
-->

## Data development

Overview data for individual development

> Metadata:
> -   Etiqueta: a ID code of the individual
> -   Imersion to Adulto: data stage was first recorded
>     -   Note data in Imersion to Adult represent dates and must be
>         converted properly!
> -   SP: species - Sexo: Sex - Temp: Temperature individual developed
> -   Locality
> -   ID: Is a full ID include etiqueta, temperature, species and
>     locality
> -   Days to change ontogenetic stage
>     -   L2_L1 means, Become stage L2 from stage L1
>     -   L3_L2 means, Become stage L3 from stage L2
>     -   Adulto_Imersion represent the total number of days from first
>         immersion to adult stage
>     -   So on...

```{r}
dados<-read.xlsx("Book1.xlsx",sheet="Goias_CLEAN_2")
  dados<-dados%>%select(-L1_IMERSION)

dados
```

## Data eclosion

```{r}
dados_eclosion<-read.xlsx("Experimento 1 (Temperatura).xlsx",sheet="Ovitrampa Goiás ") #Tem um "espaço a mais na aba ovitrampa Goiais, cuidado!

dados_eclosion 
```

# Eclosion model

## Data preparation

```{r}
dados_eclosion$Temperatura<-c(25,30,30,20,20,20,20,15,15,15,15)	#Necessário para tirar o "ºC"

dados_eclosion$Temperatura<-as.numeric(dados_eclosion$Temperatura)

dados_eclosion%>%
group_by(Temperatura)%>%
summarise(Eclosion_rate=mean((Ovos.eclodidos/Total.de.ovos)*100),
sd=sd((Ovos.eclodidos/Total.de.ovos)*100),
N=n())
```

## Model implementation

### Full model and model selection

```{r}
eclosion_mod<-glm(cbind(Ovos.eclodidos,Total.de.ovos)~
              Temperatura+I(Temperatura^2),family = binomial("logit"),data=dados_eclosion)

dredge(eclosion_mod)
```

### Best Model validation

Best models is the full model so no additional script is necessary

```{r}
simulateResiduals(eclosion_mod, plot =F)%>%testResiduals()
```

## Model validation
```{r}
simulateResiduals(eclosion_mod, plot =F)%>%testResiduals()

```


## Model inference

Now, use the best model to drawn conclusions

### Summary

```{r}
summary(eclosion_mod)
```

### Optimum temperature

Values predicted by best model was extract for temperature from 20ºC to
30ºC by 0.5ºC intervals x is x-axis value, it means, temperature in this
case

Optimum temperature is

```{r}
eclosion_mod_pred<-ggpredict(eclosion_mod, terms = c("Temperatura[20:30,by=.5]"))

eclosion_mod_pred%>%as.tibble()%>%
	slice_max(predicted,n = 5)%>%arrange(x)%>%select(-group) #x is temperature

```

### Plot

```{r}
plot_model(eclosion_mod, type = "pred", terms = "Temperatura[15:30]",show.data = T,colors = "bw")+
xlab("Temperature")+ylab("Egg hatching probability")+
labs(title=NULL)+
theme_bw(base_size=16)
```

# Eggs survival

<!--# Curiosamente, apesar do baixo N amostral, o modelo apresentou uma boa performance. Mesmo assim não será utilizado pois não incorpora novas informações que jã não possam ser retiradas do modelo de eclosão-->

## Data preparation

```{r}
surv <- data.frame(
  Vivos = c(7, 109, 80),
  Mortos = c(3, 19, 14),
  Temperatura = c(20, 25, 30))%>%
mutate(Total=Vivos+Mortos)

surv
```

## Model implementation and validation

A simple glm with binomial distribution was used

DHARMa package was used because there is very few observation and the
`DHARMa::imulateResiduals` function improves performance in this case

```{r}
mod_surv<-glm(cbind(Mortos,Total)~
        Temperatura+I(Temperatura^2),family = binomial("logit"),data=surv)

dredge(mod_surv)

# Modelo validation
simulateResiduals(mod_surv, plot =F)%>%testResiduals()

```

## Model inference

```{r}
mod_surv%>%parameters::parameters()


plot_model(mod_surv, type = "pred", terms = "Temperatura[20:30]",show.data = T,colors = "bw")+
xlab("Temperature")+ylab("Survival probability")+
labs(title=NULL)+
theme_bw(base_size=16)

```

# Development

## Data preparation

```{r}
dados_long<-dados%>%
	select(SEXO,temp,L2_L1:ADULTO_IMERSION)%>%
		pivot_longer(!c(SEXO,temp),
			names_to="Transition",values_to="Days")%>%
filter(Transition!="ADULTO_IMERSION")	# Need to be removed to avoid being double accounted

dados_long
```

### Data summary and standardization

```{r}
dados%>%
	select(SEXO,temp,L2_L1:ADULTO_IMERSION)%>%
		pivot_longer(!c(SEXO,temp),
			names_to="Transition",values_to="Days")%>%
#	group_by(SEXO,Transition,temp)%>%	#Use this for a complete summary
	group_by(temp)%>%filter(Transition=="ADULTO_IMERSION")%>%	
		summarise(Mean_Days=round(mean(Days),2),
			SD_Days=round(sd(Days),2),
			Min=min(Days),
			Max=max(Days),
			low95=quantile(Days,0.025),
			high95=quantile(Days,0.975))%>%
							data.frame()

```

### Data transformation

Necessary step to reduce dispersion and make ontogenic stages comparable



```{r}
# Log transformation to reduce dispersion
dados_log<-dados %>%
  mutate(SEXO=as.factor(SEXO))%>%  
  mutate(across(L2_L1:ADULTO_IMERSION,log))

#z-score transformation to make models among different stages comparable
dados_std<-dados %>%
  mutate(SEXO=as.factor(SEXO))%>%  
  mutate(across(L2_L1:ADULTO_IMERSION,scale))


dados_std
dados_log

```


## Model implementation and model selection


```{r}

#----------------------------------------------------------
#	Global model
#----------------------------------------------------------
# Fit the full model
model_full<-glm(Days~temp+SEXO*Transition+I(temp^2)*SEXO*Transition,family="poisson",data=dados_long)

dredge(model_full)%>%head(n=5) 

#Select best model
# Extract best model according to AIC
best_full<-model_full%>%
	dredge%>%
		get.models(., subset = 1)


### Best model validation & summary

bestfull_glm<-glm(formula(best_full[[1]]),data=dados_long)

bestfull_glm%>%anova() #Check what covariates are significantly different than 0


plot_model(bestfull_glm, type = "pred", terms = c("temp[20:30]","Transition"),show.data = F)


#### Full model did not pass model assumptions
#Uniformity: not passed
#Dispersion: passed
#Outliers: not passed


simulateResiduals(bestfull_glm, plot =F)%>%testResiduals()


```



## Model immersion to adult

Assess how temperature and sex driven the time to adulthood from first
time immersion

<!--
Prestar atenção:
best_alldev include all variables (the global model).
Os modelos construídos para cada estágio foram incluídos no objeto mod_trans
-->


```{r}
#	MODEL - Full development
modeloADULTO_IMERSAO <- glm(ADULTO_IMERSION ~ SEXO+temp+I(temp^2), data = dados_log)
```


### Model selection and validation

```{r}
#Multi model selection
dredge_out<-dredge(modeloADULTO_IMERSAO)
dredge_out
```

```{r}
best_alldev<-get.models(dredge_out, subset = 1)
best_alldev[[1]]%>%parameters::parameters(exponentiate=FALSE) 
best_alldev[[1]]%>%anova()
```

```{r}
# Effect of the sex 
(best_alldev[[1]]%>%parameters::parameters(exponentiate =TRUE))[2,]
```

<!--"exponentiate =TRUE" Effect of the sex expoential for better biological meaning --> 


<!-- 
#Warning message "Context: nKcase - wrong family"
#This message appear because values still looking like discretized after log and poisson 
-->

```{r message=TRUE, warning=TRUE}
simulateResiduals(best_alldev[[1]],plot=T)
```


Ainda que o modelo não tenha se adequado a todos os pressupostos, essa diferença parece ser mínima e o modelo parece ter apresentado um bom ajuste quando olhamos os resíduos dos valores ajustados 


```{r}
hist(best_alldev[[1]]$residuals,xlab="Residuals",main=NULL)
```

```{r}
best_alldev_glm<-best_alldev[[1]]
```


### Plot

```{r message=FALSE, warning=FALSE}
plot_model(best_alldev_glm, show.data = T , jitter = .3,
	type = "pred", terms = c("temp[20:30]","SEXO"),transform)+
labs(x="Temperature",
	y="log(Days)",
	color="Sex")+
#xlab("Temperature")+
#ylab("log(Days)")+
labs(title=NULL)+
theme_bw(base_size=16)+ 
  theme_bw(base_size = 16)+
#  scale_y_continuous(
#    breaks = seq(log(1),log(40),by=log(log(5))),  # Intervalos de 5 em 5 de 1 a 40
#    labels = function(x) scales::label_number()(exp(x)))+
  scale_x_continuous(
    breaks = seq(20,30,by=5))

```



### Valores preditos

```{r message=FALSE, warning=FALSE}
ggpredict(best_alldev_glm,,terms = c("temp[20,25,30]"))%>%
  data.frame()%>%
  mutate(across(c(predicted, conf.low, conf.high), exp))  


# Considering sex
best_alldev_pred<-ggpredict(best_alldev_glm,,terms = c("temp[20:30,by=.5]","SEXO"))%>%
data.frame()%>%
  mutate(across(c(predicted, conf.low, conf.high), exp)) %>%
  mutate(SE = (conf.high - conf.low) / (2 * 1.96))


best_alldev_pred
```


<!-- Não lmebro para que serve essa parte do código 
best_alldev_pred%>%
group_by(group)%>%
  mutate(
    rank_min = min_rank(predicted),  # Classifica os valores por 'predicted'
    rank_max = dense_rank(desc(predicted))  # Classifica em ordem decrescente
  ) %>%
  filter(rank_min <= 3 | rank_max <= 3) %>%  # Mantém valores com rank <= 3 em ambas as classificações (isso é, os 3 primeiros)
  ungroup()  # Desagrupa após a operação
-->


```{r message=FALSE, warning=FALSE}
ggpredict(best_alldev_glm,,terms = c("temp[20:30,by=.5]","SEXO"))%>%
data.frame()%>%
  mutate(across(c(predicted, conf.low, conf.high), exp)) %>%
  mutate(SE = (conf.high - conf.low) / (2 * 1.96))%>%
group_by(group)%>%
  slice_min(order_by = predicted, n = 3)%>%
relocate(group,x)
```
# Development by stage

## Piecewise model fitting

We created an acessory function to help validate all models with one hit

```{r}
#	PIECEWISE MODEL FITTING

Stand_residout<-function(X){
 simul<-DHARMa::simulateResiduals(X)
	temp<-DHARMa::testResiduals(simul)
 out<-data.frame(
		Var=c("Uniformity(KS(D))","Dispersion","Outliers"),
			Stat=c(as.numeric(temp$uniformity[[1]]),
				as.numeric(temp$dispersion[[2]]),
				paste(as.numeric(temp$outliers[[1]]),"of",as.numeric(temp$outliers[[2]]))),
		p.value=c(temp$uniformity[[2]],
				temp$dispersion[[5]],
				paste(temp$outliers[[3]])))
return(out)}
```

Model fitting

```{r, result='hide'}
modL2_L1 <- glm(L2_L1 ~ temp+I(temp^2)+SEXO, data = dados_std)
modL3_L2 <- glm(L3_L2 ~ temp+I(temp^2)+SEXO, data = dados_std)
modL4_L3 <- glm(L4_L3 ~ temp+I(temp^2)+SEXO, data = dados_std)
modPUPA_L4 <- glm(PUPA_L4 ~ temp+I(temp^2)+SEXO, data = dados_std)
modADULTO_PUPA <- glm(ADULTO_PUPA ~ temp+I(temp^2)+SEXO, data = dados_std)

# Include all models in a list
mod_trans<-list(
  modL2_L1,
  modL3_L2,
  modL4_L3,
  modPUPA_L4,
  modADULTO_PUPA)
```

```{r message=FALSE, warning=FALSE, result='hide'}
# Check its parameters
lapply(mod_trans,Stand_residout) # Check residuals
lapply(mod_trans,performance::r2) #Check R² for each model
lapply(mod_trans,parameters::parameters) # Extract parameters for each model
```

Two models example 


<!-- 
esses dois exemplos não foram aleatórios
 lembrar e descrever pq eles foram escolhidos
-->

```{r}
plot_model(modL2_L1 , show.data = T , jitter = .3,
	type = "pred", terms = c("temp[20:30]","SEXO"),transform)


plot_model(modL3_L2 , show.data = T , jitter = .3,
	type = "pred", terms = c("temp[20:30]","SEXO"))

```




## Estimate the effect of variables on stages

```{r}
stats<-c("Intercept","Temp","Temp_quad","SEXO")

effect_temp<-lapply(
			lapply(
			  mod_trans,
					parameters::parameters),
						data.frame,stats=stats)

names(effect_temp)<-c("L1_L2","L2_L3","L3_L4","L4_Pupa","Pupa_Adulto")

effect_temp<-effect_temp%>%bind_rows(.,.id="Stages")%>%as_tibble()%>%
select( Stages,stats,Coefficient,SE,CI,CI_low,CI_high,p)%>%
filter(stats!="Intercept")%>%
mutate(p.value=ifelse(p>=0.05,"Non-sig","sig"))

effect_temp

```


### Effect size by stage - plot

```{r}
effect_temp%>%
filter(stats!="Intercept")%>%
mutate(stats=ifelse(stats=="SEXO","diff F/M",stats))%>%
#filter(Stages!="Adulto_imersion")%>%
mutate(Stages=factor(Stages,levels=unique(effect_temp$Stages)))%>%
ggplot(.,aes(x=Stages,y=Coefficient,group=stats))+
#geom_line()+
geom_pointrange(aes(ymin = CI_low, ymax = CI_high,color=p.value))+
facet_wrap(.~stats,scale="free_y")+
theme_bw(base_size=16)+
geom_hline(yintercept=0, linetype="dashed", color = "red",linesize=.5)+
scale_color_manual(values=c("grey70","black"))+
ylab("Estimated effect \non development")+xlab(NULL)+
#ylab(paste0("Estimated effect of temperature \n on development ",bquote(beta[max](mu~mol ~CO[2]~ m^-2~s^-1))))+
#ylab(expression(atop("Estimated effect of temperature", "on development ( -1* "~beta[Temp]~")"))) +
#theme(axis.text.x = element_text(angle = 75, vjust = 1, hjust=1))
theme(axis.text.x = element_text(angle = 90, vjust = .5, hjust=1))+
facet_wrap(.~stats,scales="free_y")
```

